### 2020-12-15 Paper/Paper.Rmd
Just putting the titles

---
title: 'A reproduction of Newspapers in Times of Low Advertising Revenues'
author: "Kar Hian Ong"
thanks:  "Code and data are available at: https://github.com/karhian/s4-R-5"
date: "15/12/2020"
abstract: | 
  | 
  | **Keywords:** Newspapers, least squares, Advertising, Internet
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

With the rise of the internet, the purchase of newspaper is in the decline.
In this paper I attempt to reproduce Newspapers in Times of Low Advertising Revenues by Charles Angelucci and Julia Cagé

# Data

# Model

# Discussion

## Weakness

# References